---
title: "Assignment 1"
author: "Chinmaya Kumar Bansal"
date: "4 September 2018"
header-includes:
   - \usepackage{bbm}
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
  includes:
    in_header: check1.sty
---
###Question 1:
####Consider the probability density function p(x) =(c/x^4) for x>=1,where c is a constant. Generate 100 random samples from this distribution and plot a histogram. How close is the average of the samples to the expected value of X?
### Answer 1:

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r equation1, echo=FALSE}
x=runif(100,min=0,max=1)
a=floor(1/x)
p=3/(a**2)
```

```{r graph1 , echo=FALSE}
b=hist(p,main = "Histogram of Samples",xlab = "Value of Samples")
#plot(b)
```

```{r average1,echo=FALSE}
avg=mean(p)
```

```{r expected1 value, echo =FALSE}
f=function(a){
  3/a**3
}
c=integrate(f,1,Inf)
```
Average value of samples is
```{r print avg1, echo=FALSE}
print(avg)
```
Expected value of X is
```{r print expected1, echo=FALSE}
print(c)
```


\pagebreak


###Question 2:
####Draw a 2-D plot in which the Y-axis represents V(d), the volume of a d-dimensional unit ball, and the X-axis represents d= 1,2,3, . . .. State your observations
### Answer 2:

```{r work2, echo=FALSE}
x=(1:20)
y=(pi^(x/2))/(gamma((x/2)+1))
```

```{r plot2,echo=FALSE}
plot(x,y,ylab = "Volume", xlab = "Number of dimensions",type = "b",main = "Volume of d-dimensional unit ball")
```
\pagebreak

###Question 3:
####Draw a 2-D plot in which the Y-axis represents S(d), the surface area of a d-dimensional unit ball, and the X-axis represents d= 1,2,3, . . .. State your observations.

###Anser 3:

```{r code3 , echo=FALSE}
x=(1:20)
y=(2*pi^((x+1)/2))/(gamma((x+1)/2))
plot(x,y,ylab = "Surface Area", xlab = "Number of dimensions",type = "b", main = "Surface Area of d-dimensional unit ball")
```

\pagebreak

###Question 4:

####Draw a 3-D plot in which the Z-axis represents V(d), the volume of a d-dimensional ball of radius R, and the X-axis represents d, and the Y-axis represents the radius R. State your observations.

###Answer 4:
```{r work4, echo=FALSE}
x=(1:10)#dimension
y=(1:10)#radius
z=function(x,y){((pi^(x/2))*(y**x))/(gamma((x/2)+1))}
```

```{r plot4,echo=FALSE}
z=outer(x,y,z)
persp(x,y,z,xlab = "Number of Dimensions - X Axis",ylab = "Radius - Y Axis", zlab = "Volume of Sphere - Z Axis",main = "Plot of Volume ", theta = 30,col = "green",phi=0,expand = 0.5)
```

\pagebreak

###Question 5:

####Draw a 3-D plot in which the Z-axis represents S(d), the surface area of a d-dimensional ball of radius R, and the X -axis represents d, and the Y-axis represents the radius R. State your observations.
###Answer 5:
```{r work5, echo=FALSE}
x=(1:10)#dimension
y=(1:10)#radius
z=function(x,y){(2*(pi^((x+1)/2))*(y**x))/(gamma((x+1)/2))}
```

```{r plot5,echo=FALSE}
z=outer(x,y,z)
persp(x,y,z,xlab = "Number of Dimensions - X Axis",ylab = "Radius - Y Axis", zlab = "Surface Area of Sphere - Z Axis",main = "Plot of Surface Area ", theta = 30,col = "green",phi=0,expand = 0.5)
```



\pagebreak

###Question 6:
####Generate 20 points uniformly at random on a 900 -dimensional sphere of radius 30. Calculate the distance between each pair of points. Then, select a method of projection and project the data onto subspaces of dimension k = 100, 50,10,5,4,3,2,1 and calculate the difference between k^(1/2) times the original distances and the new pair-wise distances. For each value of k what is the maximum difference as a percent of k^(1/2).

###Answer 6: 

```{r sphere6, echo = FALSE}
library(hitandrun)

sample <- hypersphere.sample(900, 20)       #Generating 20 Random points
fin_sample <- 30*sample                     #Radius to 30
```

```{r distance6, echo = FALSE}
d <- dist(fin_sample,method = "euclidean")  #Distance between points
```

```{r function6, echo = FALSE}
  scaling <- function(d,ndim){              #Scaling to n-deimension
  fit <- cmdscale(d, eig = FALSE, k = ndim) #Classical Multidimensional Scaling
  d_new <- dist(fit)                        #Calculating new distances  
  diff <- (((k^0.5)*d) - d_new)             #Calculating the differnce
  print(paste("Subspace of dimension:",k))
  print(max(diff))
}
```

The differences in euclidean distances for the subspace projection is as follows:
```{r 1dim,echo = FALSE}
k = 1
scaling(d,k)
```

```{r 2dim,echo = FALSE}
k = 2
scaling(d,k)
```

```{r 3dim,echo = FALSE}
k = 3
scaling(d,k)
```

```{r 4dim,echo = FALSE}
k = 4
scaling(d,k)
```

```{r 5dim,echo = FALSE}
k = 5
scaling(d,k)
```

```{r 10dim,echo = FALSE}
k = 10
scaling(d,k)
```
